---
title: "Reproducible Research - Assignment 1"
output: html_document
date: "January 2015"
---

First we need to load the data. Let's look at the first few lines.
```{r}
activity <- read.csv('activity.csv')
head(activity)
```


Now we want to sum up the total number of steps taken each day.
```{r}
activitySum <- aggregate(activity$steps, FUN = sum, by = list(activity$date))
```


What is the mean total number of steps taken per day?
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
qplot(x, data = activitySum, xlab = "Steps taken daily")
summary(activitySum)
meanx <- mean(activitySum$x, na.rm = TRUE)
medianx <- median(activitySum$x, na.rm = TRUE)

```
The mean is `r meanx` steps and the median is `r medianx` steps per day


```{r}
library(plyr)
library(ggplot2)
dayavg <- ddply(activity, .(interval), summarize, mean = mean(steps, na.rm = TRUE))
qplot(interval, mean, data = dayavg, geom = 'line',
      xlab = "Time", ylab = "Avg number of steps")
maxinterval <- dayavg[which(dayavg$mean == max(dayavg$mean)), 1]
```
The greatest number of steps occurs in interval `r maxinterval`


Imputing Missing Values
```{r}
ok <- complete.cases(activity)
missing <- sum(!ok)
```
There are `r missing` rows containing missing values (NA's)

```{r}
lookupInterval <- function(interval){
  avgSteps <- dayavg[match(interval, dayavg$interval),2]
  avgSteps
}

cleanActivity <- activity
cleanActivity[is.na(cleanActivity),][,1] <-
  lookupInterval(cleanActivity[is.na(cleanActivity),][,3])

cleanActivitySum <- aggregate(cleanActivity$steps, FUN = sum, by = list(cleanActivity$date))
qplot(x, data = cleanActivitySum, xlab = "Steps taken daily")
cleanMean <- mean(cleanActivitySum$x, na.rm = TRUE)
cleanMedian <- median(cleanActivitySum$x, na.rm = TRUE)
```
The mean is `r cleanMean` steps and the median is `r cleanMedian` steps per day.  
After imputing values, the mean and median are now equal.  

For days that are not weekends, more steps are taken earlier in the morning.
```{r}
cleanActivity$day <- as.factor(weekdays(as.Date(cleanActivity$date)))
cleanActivity$weekend <- cleanActivity$day %in% c("Saturday", "Sunday")

cleanAvg <- ddply(cleanActivity, .(interval, weekend),
                    summarize, mean = mean(steps, na.rm = TRUE))

qplot(interval, mean, data = cleanAvg, geom = 'line', facets = weekend~.,
      xlab = "Time", ylab = "Avg number of steps")
```



